---
title: "Practice Fixed Effect and Random Effects Meta-analysis"
author: "ks"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Attach libraries.

```{r packages}
for (pkg in c("tidyverse","MASS", "readxl","dplyr","meta","metafor","netmeta")) 
   library(pkg, character.only = TRUE)
```

## Simulation function

I will simulate binary data using the following design:   

** 8 Studies  
** n1 = n2 in each Study     
** n1 ~ Negative binomial(mu,mu+mu^2/theta); mu=80,theta=5  
** p0 = 0.3    
** or = 2.0

Sample sizes vary from study-to-study following a Negative binomial distribution. What does the distribution look like?
```{r distn-negbin}
set.seed(20)
n1 <- rnegbin(1000,mu=80,theta=5)
hist(n1,main=paste("mean=",round(mean(n1),0)))
```

Write a function to compute probability, p, from the odds(p).
```{r myfunc}
invodds <- function(x){
    p <- x/(1+x)
}
```

Write a function to simulate data from 1 data set.
```{r simdat-and-sum}
study.sum <- function(n1,p1=0.3,or=2.0) {
        gp  <- factor(rep(c("group1", "group2"), each=n1))
        # odds(p2) = or*p1/(1-p1)
        p2  <- invodds(or*p1/(1-p1))

      yvec <- rbinom(n = n1*2, size = 1, 
                    prob = rep(c(p1,p2),each=n1))
      tb <- table(yvec,gp)
       a <- tb[2,2]
       b <- tb[2,1]
       c <- tb[1,2]
       d <- tb[1,1]
       es <- a*d/b/c
       lnes <- log(es)
       selnes <- sqrt(1/a + 1/b + 1/c + 1/d)
       te <- a
       tn <- n1
       ce <- b
       cn <- n1
       
      study.out <- c(te,tn,ce,cn,es,lnes,selnes)
      study.out
}
```

## Common OR=2

Now let's simulate 8 data sets each with True OR = 2.0.
```{r simdat-commonor}
nstudy <- 8
cosumdat <- matrix(NA,nrow=nstudy,ncol=7)
colnames(cosumdat) <- c("te","tn","ce","cn","es","lnes","selnes")
set.seed(91)
n1vec <- rnegbin(nstudy,mu=80,theta=5)
for (i in 1:nstudy) {
    cosumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=2.0)
}
cosumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),cosumdat)
cosumdat
```


We can analyze the raw counts using the function metabin().  We'll use the data like this:

* te = Treatment group, number of events
* tn = Treatment group, number of observations
* ce = Control group, number of events
* cn = Control group, number of observations

What do you expect these results to show and why?

```{r metabin}
mb1 <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=cosumdat)
forest(mb1)
```

Hmmmmmm. The fixed effect and random effects summary is the same. Why?


##  ln(OR) ~ N(log(2),0.8)

Now let's model the ln(OR) ~ N(log(2),0.8).

First let's see what this distribution gives us.
```{r dist-lnor}
set.seed(20)
simlnor <- rnorm(10000,log(2),0.8)
hist(simlnor,main=paste("mean=",round(mean(simlnor),3)))
```

```{r simdat-manytheta}
nstudy <- 8
rosumdat <- matrix(NA,nrow=nstudy,ncol=7)
colnames(rosumdat) <- c("te","tn","ce","cn","es","lnes","selnes")
#set.seed(98)
set.seed(20)
simor <- exp(rnorm(nstudy,log(2),0.8))
for (i in 1:nstudy) {
    rosumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=simor[i])
}
rosumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),rosumdat)
rosumdat
```

```{r plot-true-est-ors}

```

```{r metabin-re}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=rosumdat)
forest(mb)
```

What is the summary OR estimate (95% CI)? Which model result should be reported and why?  

## More Studies

What happens if we had 10xs the number of studies?
How does this affect the summary OR estimate (95% CI) under the common OR model?
```{r simdat-commonor-morestudies}
nstudy <- 8*10
set.seed(99)
n1vec <- rnegbin(nstudy,mu=50,theta=5)
sumdat <- matrix(NA,nrow=nstudy,ncol=7)
colnames(sumdat) <- c("te","tn","ce","cn","es","lnes","selnes")
for (i in 1:nstudy) {
    sumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=2.0)
}
sumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),sumdat)
sumdat
```

```{r metabin5}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=sumdat)
summary(mb)
```

```{r simdat-manyor-morestudies}
# same as above
#nstudy <- 8*25
#set.seed(99)
#n1vec <- rnegbin(nstudy,mu=50,theta=5)
 
sumdat <- matrix(NA,nrow=nstudy,ncol=7)
set.seed(20)
colnames(sumdat) <- c("te","tn","ce","cn","es","lnes","selnes")
simor <- exp(rnorm(nstudy,log(2),0.8))
for (i in 1:nstudy) {
    sumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=simor[i])
}
sumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),sumdat)
sumdat
```

```{r metabin6}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=sumdat)
summary(mb)
```


```{r sessionInfo}
sessionInfo()
```

