---
title: "Class activity on Heterogeneity"
author: "ks"
date: "3/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Libraries and Functions

```{r packages}
for (pkg in c("tidyverse","MASS", "readxl","dplyr","meta","metafor","netmeta")) 
   library(pkg, character.only = TRUE)
```

Sample sizes vary from study-to-study following a Negative binomial distribution. 
```{r distn-negbin}
set.seed(20)
n1 <- rnegbin(1000,mu=80,theta=5)
hist(n1,main=paste("mean=",round(mean(n1),0)))
```

Write a function to compute probability, p, from the odds(p).
```{r myfunc}
invodds <- function(x){
    p <- x/(1+x)
}
```

Write a function to simulate data from 1 data set.
```{r simdat-and-sum}
study.sum <- function(n1,p1=0.3,or=2.0) {
      # odds(p2) = or*p1/(1-p1)
      p2  <- invodds(or*p1/(1-p1))

      yvec <- rbinom(n = 2, size = n1, 
                    prob = c(p1,p2))
      study.out <- c(te=yvec[2],
                     tn=n1,
                     ce=yvec[1],
                     cn=n1)
      study.out
}
```

## Simulate Data: small $\tau$

I will simulate binary data using the following design:   

** n1 = n2 in each Study     
** n1 ~ Negative binomial(mu,mu+mu^2/theta); mu=80,theta=5  
** p0 = 0.3    
** ln(OR) ~ N(log(2),0.4).

First let's see what this distribution ln(OR) ~ N(log(2),0.4) gives us.
```{r dist-lnor}
set.seed(20)
simlnor <- rnorm(10000,log(2),0.4)
hist(simlnor,main=paste("mean=",round(mean(simlnor),3)))
```

Simulate 100 datasets with random ORs, with logOR~N(log(2),0.4). We will sample from these to do our analysis later.
```{r simdat-randomor}
nstudy <- 100
rosumdat <- matrix(NA,nrow=nstudy,ncol=4)
colnames(rosumdat) <- c("te","tn","ce","cn")
set.seed(91)
n1vec <- rnegbin(nstudy,mu=80,theta=5)

# same 100 sample sizes as above
set.seed(20)
simor <- exp(rnorm(nstudy,log(2),0.4))
for (i in 1:nstudy) {
    rosumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=simor[i])
}
rosumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),rosumdat)
rosumdat
```

## small $\tau$; N = 8 studies

```{r simdat-n8}
set.seed(5)
idx <- sample(100,8)
rosumdat.8s <- rosumdat[idx,]
rosumdat.8s
```

We can use inverse variance weights to estimate tau.
```{r metabin-DL8}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=rosumdat.8s)
summary(mb)
```

What is the summary OR estimate (95% CI)? Which model result should be reported and why?  

There are many methods for estimating tau.  Let's try REML.

```{r metabin-reml8}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau ="REML", data=rosumdat.8s)
summary(mb)
```


```{r metabin-eb8}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau="EB", data=rosumdat.8s)
summary(mb)
```

## small $\tau$; N = 80 studies

What happens if we had 10xs the number of studies?

```{r simdat-n80}
set.seed(5)
idx <- sample(100,80)
rosumdat.80s <- rosumdat[idx,]
#rosumdat.80s
```

```{r metabin-DL80}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=rosumdat.80s)
summary(mb)
```


```{r metabin6-reml80}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau="REML", data=rosumdat.80s)
summary(mb)
```


```{r metabin-eb80}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau="EB", data=rosumdat.80s)
summary(mb)
```

Same CI, different estimate of tau and tau^2.

How do these results compare to the analysis of just 8 studies?

## Simulate data: big $\tau$

First let's see what this distribution ln(OR) ~ N(log(2),0.8) gives us.
```{r dist-lnor-bt}
set.seed(20)
simlnor <- rnorm(10000,log(2),0.8)
hist(simlnor,main=paste("mean=",round(mean(simlnor),3)))
```

Simulate 100 datasets with random ORs, with logOR~N(log(2),0.8). We will sample from these to do our analysis later.
```{r simdat-bigtau}
nstudy <- 100
rosumdat <- matrix(NA,nrow=nstudy,ncol=4)
colnames(rosumdat) <- c("te","tn","ce","cn")
set.seed(91)
n1vec <- rnegbin(nstudy,mu=80,theta=5)

# same 100 sample sizes as above
set.seed(20)
simor <- exp(rnorm(nstudy,log(2),0.8))
for (i in 1:nstudy) {
    rosumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=simor[i])
}
rosumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),rosumdat)
rosumdat
```

##  big $\tau$; N = 8 studies

```{r simdat-btn8}
set.seed(5)
idx <- sample(100,8)
rosumdat.8s <- rosumdat[idx,]
rosumdat.8s
```

We can use inverse variance weights to estimate tau.
```{r metabin-btDL8}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=rosumdat.8s)
summary(mb)
```


Now let's estimate tau using REML.

```{r metabin-btreml8}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau ="REML", data=rosumdat.8s)
summary(mb)
```

Now let's estimate tau using empirical Bayes.
```{r metabin-bteb8}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau="EB", data=rosumdat.8s)
summary(mb)
```

## big $\tau$; N = 80 studies

What happens if we had 10xs the number of studies?

```{r simdat-btn80}
set.seed(5)
idx <- sample(100,80)
rosumdat.80s <- rosumdat[idx,]
#rosumdat.80s
```

```{r metabin-btDL80}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=rosumdat.80s)
summary(mb)
```


```{r metabin6-btreml80}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau="REML", data=rosumdat.80s)
summary(mb)
```


```{r metabin-bteb80}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method.tau="EB", data=rosumdat.80s)
summary(mb)
```

How does the estimate of tau (95% CI) differ using these different estimators and across these different study designs?

## SessionInfo

```{r sessionInfo}
sessionInfo()
```

