---
title: "Practice Fixed Effect and Random Effects Meta-analysis"
author: "ks"
date: "2/8/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Attach libraries.

```{r packages}
for (pkg in c("tidyverse","MASS", "readxl","dplyr","meta","metafor","netmeta")) 
   library(pkg, character.only = TRUE)
```

## Simulation function

I will simulate binary data using the following design:   

** n1 = n2 in each Study     
** n1 ~ Negative binomial(mu,mu+mu^2/theta); mu=80,theta=5  
** p0 = 0.3    
** or = 2.0

Sample sizes vary from study-to-study following a Negative binomial distribution. What does the distribution look like?
```{r distn-negbin}
set.seed(20)
n1 <- rnegbin(1000,mu=80,theta=5)
hist(n1,main=paste("mean=",round(mean(n1),0)))
```

Write a function to compute probability, p, from the odds(p).
```{r myfunc}
invodds <- function(x){
    p <- x/(1+x)
}
```

Write a function to simulate data from 1 data set.
```{r simdat-and-sum}
study.sum <- function(n1,p1=0.3,or=2.0) {
      # odds(p2) = or*p1/(1-p1)
      p2  <- invodds(or*p1/(1-p1))

      yvec <- rbinom(n = 2, size = n1, 
                    prob = c(p1,p2))
      study.out <- c(te=yvec[2],
                     tn=n1,
                     ce=yvec[1],
                     cn=n1)
      study.out
}
```


Simulate 100 data sets with a True OR of 2.0.
```{r simdat-commonor}
nstudy <- 100
cosumdat <- matrix(NA,nrow=nstudy,ncol=4)
colnames(cosumdat) <- c("te","tn","ce","cn")
set.seed(91)
n1vec <- rnegbin(nstudy,mu=80,theta=5)
for (i in 1:nstudy) {
    cosumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=2.0)
}
cosumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),cosumdat)
cosumdat
```

Now let's model the ln(OR) ~ N(log(2),0.8).

First let's see what this distribution gives us.
```{r dist-lnor}
set.seed(20)
simlnor <- rnorm(10000,log(2),0.8)
hist(simlnor,main=paste("mean=",round(mean(simlnor),3)))
```

Simulate 100 datasets with random ORs, with logOR~N(log(2),0.8).
```{r simdat-randomor}
nstudy <- 100
rosumdat <- matrix(NA,nrow=nstudy,ncol=4)
colnames(rosumdat) <- c("te","tn","ce","cn")
# same 100 sample sizes as above
set.seed(20)
simor <- exp(rnorm(nstudy,log(2),0.8))
for (i in 1:nstudy) {
    rosumdat[i,] <- study.sum(n1=n1vec[i],p1=0.3,or=simor[i])
}
rosumdat <- cbind.data.frame(study = paste0("Study",1:nstudy),rosumdat)
rosumdat
```

## Common OR=2

Now let's sample 8 data sets each with True OR = 2.0 and do a meta-analysis.
```{r sample-8studies-commonor}
set.seed(5)
idx <- sample(100,8)
cosumdat.8s <- cosumdat[idx,]
cosumdat.8s
```


We can analyze the raw counts using the function metabin().  We'll use the data like this:

* te = Treatment group, number of events
* tn = Treatment group, number of observations
* ce = Control group, number of events
* cn = Control group, number of observations

What do you expect these results to show and why?

```{r metabin}
mb1 <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=cosumdat.8s)
forest(mb1)
```

Which summary OR 95% CI is larger and why?


##  ln(OR) ~ N(log(2),0.8)

Now let's model the ln(OR) ~ N(log(2),0.8).

```{r simdat-manytheta}
#set.seed(5)
#idx <- sample(100,8)
rosumdat.8s <- rosumdat[idx,]
rosumdat.8s
```

```{r plot-true-ors}
par(mfrow=c(1,2))
plot(log(simor[idx]),1:8,xlab = "true lnOR",
     ylab="study",pch=16)
plot(simor[idx],1:8,xlab = "true OR",
     ylab="study",pch=16)
```

```{r metabin-re}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=rosumdat.8s)
forest(mb)
```

What is the summary OR estimate (95% CI)? Which model result should be reported and why?  

## More Studies

What happens if we had 10xs the number of studies?

Let's first consider having 1 true OR = 2.0.
```{r simdat-commonor-morestudies}
set.seed(5)
idx <- sample(100,80)
cosumdat.80s <- cosumdat[idx,]
cosumdat.80s
```

```{r metabin5}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=cosumdat.80s)
summary(mb)
```

How do these results compare to the analysis of just 8 studies?

Now we'll revisit the model with lnOR~Normal distribution.
```{r simdat-manyor-morestudies}
set.seed(5)
idx <- sample(100,80)
rosumdat.80s <- rosumdat[idx,]
rosumdat.80s
```

```{r metabin6}
mb <- metabin(event.e=te, n.e=tn, 
               event.c=ce, n.c=cn, studlab=study, sm="OR", 
               method="I", data=rosumdat.80s)
summary(mb)
```

How do these results compare to the analysis of just 8 studies?

## SessionInfo

```{r sessionInfo}
sessionInfo()
```

