---
title: "Example of Network Meta Analysis"
author: "ks"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Dataset

```{r libraries}
if(!require(netmeta)) install.packages("netmeta")
library(netmeta)
```

Load dataset:
```{r data}
data(Senn2013)
data15 <- Senn2013
data15
```

```{r variable-names}
names(data15)
```


How many drug comparisons?
```{r nte}
nrow(data15)
```

How many unique studies?
```{r nstudies}
length(unique(data15$studlab))
```

If we have more comparisons than studies, not all comparisons are independent and our network analysis needs to take into account the correlation between treatment comparisons that originate from a single study.  We do this by adjusting the standard errors for the treatment comparisons originating from a single study (make them bigger).

## Network meta-analysis

```{r netmeta}
mn1 <- netmeta(TE, seTE, treat1, treat2, studlab,
         data=data15, sm="MD")
```


```{r graph-it}
netgraph(mn1, seq=c("plac", "benf", "migl", "acar", "sulf",
                     "metf", "rosi", "piog", "sita", "vild"))
```

Summarize output:

```{r summary}
print(summary(mn1))
```

## More graphs

Setting multiarm = TRUE  will color treatments from multi-arm studies.
```{r  moregraphs-multiarm}
netgraph(mn1, start="random", iterate=TRUE,
 col="darkgray", cex=1.5, multiarm=TRUE,
 points=TRUE, col.points="green", cex.points=3)
```

## Placebo Reference

```{r placebo-ref}
summary(mn1, ref="plac")
```

Forest Plot for fixed effect model:

```{r placebo-ref-forest}
forest(mn1, ref="plac",
       xlab="HbA1c difference",
       pooled="fixed")
```

Forest Plot for random effects model:

```{r placebo-ref-reforest}
forest(mn1, ref="plac",
       leftlabs="Contrast to Placebo",
       xlab="HbA1c difference",
       pooled="random")
```

## Heterogeneity

```{r heterogeneity}
round(decomp.design(mn1)$Q.decomp, 3)
```

## Net heatmaps

This displays for each treatment effect estimate, two types of information simultaneously. These are:  
(1) the contribution of each design to this estimate, and  
(2) the extent of inconsistency due to each design.

For the fixed effect model:

```{r netheat}
netheat(mn1)
```

Rows & columns are the within-design treatment effect estimates. The area of the grey rectangles give the %contribution to the estimate. Largest areas on the diagonal indicate direct effects carry the most weight. Note some rows/columns have several boxes with large areas (e.g. acar:placebo, acar:sulf). labels with _ indicate the 3-treatment study.

The color ranks the between-design contribution to each estimate, with highest in top left corner. So metf:sulf and rosi:sulf designs have the largest between-design heterogeneity. They also contribute inconsistent information to the treatment estimate of the other (Positions 1,2 and 2,1 are red).

For the random effects model:

```{r netheat-random}
netheat(mn1, random=TRUE)
```

Using a random effects model reduces the between-design heterogeneity. 

## Bland–Altman Plot 

Here is a plot to assess the effect of heterogeneity
on treatment effect estimates.

```{r baplot}
set.seed(125)
fe <- mn1$TE.nma.fixed
re <- mn1$TE.nma.random
par(mar=c(5,5,2,3))
plot(jitter((fe+re)/2, 5), jitter(fe-re, 5),
      xlim=c(-1.2, 1.2),
      ylim=c(-0.25, 0.25),
      xlab="Mean treatment effect (in fixed effect and random
effects model)", 
      ylab="Diff. of treat effect 
      (FE - RE model ests.)")
abline(h=0)
```

The fixed and random effects estimates are most different for the two studies on the bottom left.


## SessionInfo

```{r sessioninfo}
sessionInfo()
```