---
title: "Intro to Meta Analysis"
author: "online documentation"
date: "2/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing and Loading R Packages

We will need to install a few packages that are not in standard base r and load them.

```{r usePackages}
for (pkg in c("tidyverse", "readxl","dplyr","meta","metafor")) {
  if(!require(pkg, character.only = TRUE)) 
    install.packages(pkg)
  library(pkg, character.only = TRUE)
}
print('Done installing &/or loading libraries')
```

## Example

Now, let's read in the data from the Excel spreadsheet. This example is taken from the textbook, **Introduction to Meta Analysis**, by Borenstein, Hedges, Higgins and Rothstein, 2009. John Wiley & Sons, Ltd.

```{r ReadData}
exdata <- read_xlsx("data/Binary_Data_Example.xlsx",sheet=1)
exdata
```

* Tevents = Treatment group, number of events
* TN = Treatment group, number of observations
* Cevents = Control group, number of events
* CN = Control group, number of observations



Next, we run the fixed effects model meta analysis for binary data.

```{r FixedEffect}
mb1 <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="I", data=exdata, studlab=Study)
mb1
```


We get a (fixed effect) summary OR of 0.48 with 95\% confidence interval (CI) (0.36, 0.66)  ( p < 0.0001 ).

Instead of just the fixed effects model, we also get the summary from the random effects model.  The output tells us that it uses the DerSimonian-Laird estimator for tau^2 in the random effects model. The summary OR estimate from the random effects model is 0.57 (95\% CI: 0.36-0.91) ( p < 0.02 ). 
There are other methods we can consider for estimating the summary from the random effects model. I'll leave this discussion for later.

Now let's use a forest plot to display the summary data.
```{r ForestPlot,fig.width=10,fig.height=4}
forest(mb1)
```


Or we can simplify the figure, which allows for larger text.
```{r SimpleForestPlot,fig.width=9,fig.height=4}
forest(mb1, comb.random=FALSE, hetstat=FALSE)
```

```{r sessionInfo}
sessionInfo()
```