---
title: "Intro-to-Dose-Response"
author: "ks"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Load libraries

Install and load the Dose-response meta-analysis package.

```{r install}
if(!require("dosresmeta", character.only = TRUE)) 
    install.packages("dosresmeta")
library("dosresmeta", character.only = TRUE)
library(tidyverse)
library(meta)
```

## Case-Control Study 

This data is from a case-control study of alcohol consumption and breast cancer.

```{r ccexample}
data("cc_ex")
head(cc_ex)
```

We need to compute the variance for the logrr, and create a variable indicating the study design (cc = case-control study).

```{r ccmutate}
cc_ex <- cc_ex %>%
 mutate(
  lnl95 = log(lb),
  lnu95 = log(ub),
  selnes = (lnu95-lnl95)/3.919928,
  type = "cc"
  )
cc_ex
```

The software will use the Greenland and Longnecker covariance formula to get the covariances of the logrr from the sample sizes (n) and number of cases (case) for each pair of dose levels.

```{r cccovar}
covar.logrr(cases = case, n = n, y = logrr, v = selnes^2, 
type = "cc", data = cc_ex)
```

Next we run the dose-response meta analysis using the Greenland and Longnecker estimate of the variance-covariance matrix.

```{r ccex-lm}
lin_le <- dosresmeta(logrr ~ dose, se = selnes, type = type, cases = case, n = n, data = cc_ex)
summary(lin_le)
```

```{r predictcc}
predict(lin_le, newdata = data.frame(dose = c(0,1,5)), expo = TRUE)
```

This says the trend OR = 1.05 (g/day), with 95\% CI: 1.00 - 1.09.

## Incidence Rate Data

This example studies the association between fiber intake (g/d) and coronary heart disease (CHD). The study compares incidence of CHD using person-years.

```{r irexample}
data("ir_ex")
head(ir_ex)
```

This time we only need to create a variable indicating the study design (ir = incidence rate study).

```{r irmutate}
ir_ex <- ir_ex %>% 
  mutate(
  type = "ir"
  )
```

Using type = "ir", the software will use the incidence rate covariance formula to get the covariances of the logrr from the sample sizes (n) and number of cases (case) for each pair of dose levels.

```{r ircovar}
covar.logrr(cases = cases, n = n, y = logrr, v = se^2, 
type = "ir", data = ir_ex)
```

Next we run the dose-response meta analysis using the estimate of the variance-covariance matrix for an incidence rate (type = ir) study.

```{r irdrmeta}
lin_le <- dosresmeta(logrr ~ dose, se = se, type = type, cases = cases, n = n, data = ir_ex)
summary(lin_le)
```

```{r irpredict}
predict(lin_le, newdata = data.frame(dose = c(0,5,10)), expo = TRUE)
```

## Cumulative Incidence Rate

Example studying high-fat dairy food (in servings/day) and colorectal cancer risk.

```{r ciexample}
data("ci_ex")
head(ci_ex)
```

Again, we create a variable indicating the study design (ci = cummulative incidence study).

```{r cimutate}
ci_ex <- ci_ex %>% 
  mutate(
  type = "ci"
  )
```

The software will use the covariance formula for cumulative incidence ('ci') data to get the covariances of the logrr from the sample sizes (n) and number of cases (case) for each pair of dose levels.

```{r cicovar}
covar.logrr(cases = cases, n = n, y = logrr, v = se^2, 
type = "ci", data = ci_ex)
```

Next we run the dose-response meta analysis using this estimate of the variance-covariance matrix.

```{r cidrmeta}
lin_le <- dosresmeta(logrr ~ dose, se = se, type = type, cases = cases, n = n, data = ci_ex)
summary(lin_le)
```

```{r cipredict}
predict(lin_le, newdata = data.frame(dose = c(0:4)), expo = TRUE)
```

This says the trend OR = 0.93, with 95\% CI: 0.89 - 0.97.

## Multiple studies 

This example includes data from six studies of alcohol intake and cardiovascular disease risk.

```{r alccvd}
data("alcohol_cvd")
alcohol_cvd
```

What types of studies are they?

Let's carry out the dose-response meta-analysis.
```{r multstudieslintrend}
lin <- dosresmeta(logrr~dose, id=id, se=se, 
                  cases = cases, n = n, 
               type = type, data = alcohol_cvd)
summary(lin)
```

Here is the summarized effect estimate.
```{r predicttrend}
predict(lin ,delta = 1, expo = TRUE)
```

What does the Q statistic test in the model above?  It tests whether there is heterogeneity in the trends estimated for the 6 studies. To show this, I'm going to estimate the trend for each study, and perform a random effects meta-analysis of the 6 estimated trend coefficients.
```{r meta-eachstudy}
studytrend <- as.data.frame(matrix(data=NA,6,3))
colnames(studytrend) <- c("Study","lnes","selnes")
studytrend[,1] <- 1:6
for (i in 1:6) {
   lin_i <- dosresmeta(logrr ~ dose, se = se, type = type, cases = cases, n = n, data = alcohol_cvd[alcohol_cvd$id==i,])
   studytrend[i,2:3] <- summary(lin_i)$coef[1:2]
}
studytrend
```

Now we do a meta-analysis of these trends.
```{r metatrend}
mb <- metagen(TE=lnes,seTE=selnes, 
               sm = "OR",backtransf = FALSE,method.tau = "REML",
               data=studytrend, studlab=Study)
summary(mb)
```

Yes! It's the same Q values as from our previous analysis of multiple studies. We calculate it by hand as follows:
```{r qtest}
sum(((studytrend$lnes - mb$TE.fixed)/studytrend$selnes)^2 )
```

Yes, again! That matches the Q-test of heterogeneity. Although the average trend is essentially flat, this statistic supports heterogeneity in the study-level trends.

## Non-linear dose-response

Now we'll fit a (non-linear) quadratic trend

```{r qt}
quadr <- dosresmeta(formula = logrr ~ dose + I(dose^2), 
                    type = type, id = id,se = se, 
                    cases = cases, n = n, data = alcohol_cvd)
summary(quadr)
```

Let's graph the results.

```{r plotresults}
with(predict(quadr, expo = TRUE, order = TRUE), {
   plot(dose, pred, log = "y", type = "l",
        ylab="ES",
        xlim = c(0, 45), ylim = c(.3, 2.4))
   lines(dose,  ci.lb, lty = 2)
   lines(dose, ci.ub, lty = 2)
   rug(dose, quiet = TRUE)
   points(alcohol_cvd$dose,exp(alcohol_cvd$logrr),pch=alcohol_cvd$id,col=alcohol_cvd$id)
})
```

## Ever/Never exposed

Let's collapse ORs by dose into an ever/never exposed. 

Let's try this for the breast cancer case-control data set.

```{r ccex}
data("cc_ex")
head(cc_ex)
```

We need to compute the variance for the logrr, and create a variable indicating the study design (cc = case-control study).

```{r ccmutate2}
cc_ex <- cc_ex %>%
 mutate(
  lnl95 = log(lb),
  lnu95 = log(ub),
  selnes = (lnu95-lnl95)/3.919928,
  ever = ifelse(dose==0,0,1),
  type = "cc"
  )
cc_ex
```


```{r everexposed}
lin_le <- dosresmeta(logrr ~ ever, se = selnes, type = type, cases = case, n = n, data = cc_ex)
predict(lin_le, newdata = data.frame(ever= c(0,1)), expo = TRUE)
```

This says the OR for breast cancer comparing alcohol drinkers vs non-drinkers  is 1.12, with 95\% CI: 0.79 - 1.59. Notice how the confidence interval contains 1.0 when we don't leverage the dose-response information.

## Session Info
```{r sessionInfo}
sessionInfo()
```

