---
title: "IUD use and Cervical Cancer Meta Analysis"
author: "K Siegmund"
date: "3/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Installing and Loading R Packages

We will need to install a few packages that are not in standard base r and load them.

```{r usePackages}
for (pkg in c("tidyverse","haven","dplyr","meta","metafor")) {
  if(!require(pkg, character.only = TRUE)) 
    install.packages(pkg)
  library(pkg, character.only = TRUE)
}
print('Done installing &/or loading libraries')
```

## Example: IUD use and Cervical Cancer

Now, let's read in the data from the paper "Intrauterine device use and cervical cancer: A systematic review and meta-analysis" by Cortessis et al. (2017).

```{r ReadData}
data <- read_dta("~kims/Google Drive File Stream/My Drive/PM605/IUD-CCstudy/Data/CCaMetaAPaper.dta")
data
```

The effect size from these studies are odds ratios. These will get analyzed on the log scale so we have to transform the data to get the log OR, and it's standard error.  The standard error for the log OR is obtained from its confidence interval. 

```{r DataPrep}
data <- data %>% 
  mutate(
  lnes = log(es),
  lnl95 = log(l95),
  lnu95 = log(u95),
  selnes = (lnu95-lnl95)/3.919928
  )
```

Next, we run both a fixed effect and random effects meta analysis in a single command.

```{r MetaA}
ma <- metagen(TE=lnes, seTE=selnes, studlab = paste(author, pub_year), sm="OR", data=data, backtransf = FALSE)
summary(ma)
```

And if we drop the option backtransf = FALSE, we get the results on the OR scale.

```{r MetaA_OR}
ma <- metagen(TE=lnes, seTE=selnes, studlab = paste(author, pub_year), sm="OR", data=data)
summary(ma)
```

Now let's use a forest plot to display the summary data. I'm going to drop the fixed effect weights from the figure.

```{r ForestPlot,fig.width=8,fig.height=6}
forest(ma, comb.fixed=FALSE)
```

Now we want to study the heterogeneity in effect estimates by study design (case-control, etc.)
```{r bystudytype}
ma <- metagen(TE=lnes, seTE=selnes, studlab = paste(author, pub_year), sm="OR", byvar = stype, comb.fixed=F, data=data)
summary(ma)
```

```{r ForestPlot2,fig.width=8,fig.height=8}
forest(ma, comb.fixed=FALSE)
```

It appears that the effect estimate does not vary depending on the desing of the study (good). Unfortunately we lost the naming of the categorical variable when we read in the data set. I will go back and fix this later.

stype 1 = Nested case-control

stype 2 = Population-based

stype 3 = Clinic-based/hospital

stype 4 = other


Now we will stratify the estimates by whether or not the study controlled for important covariates such as SES or smoking history in their analysis. 

```{r byses}
ma <- metagen(TE=lnes, seTE=selnes, studlab = paste(author, pub_year), sm="OR", byvar = ses, comb.fixed=F, data=data)
summary(ma)
```

The estimates for the studies that did and did not control for SES are rather similar (OR = 0.63 vs 0.67, p = 0.83).  Now let's look whether they controlled for smoking.

```{r bysmoking}
ma <- metagen(TE=lnes, seTE=selnes, studlab = paste(author, pub_year), sm="OR", byvar = smoking, comb.fixed = F, data=data)
summary(ma)
```

The estimates for the studies that did and did not control for smoking are rather different (OR = 0.61 vs 0.81, p=0.053).

Finally, we'll run a meta-regression to see if the OR varies as a function of the HPV rate or the Age-adjusted incidence rate.

```{r reg_hpvrate}
ma <- metagen(TE=lnes, seTE=selnes, studlab = paste(author, pub_year), sm="OR", comb.fixed = F, data=data[!is.na(data$hpvrate),])
ma.mr <- metareg(ma,hpvrate,method.tau = "REML")
print(ma.mr)
```

The OR doesn't vary as a function of the HPV rate.

```{r bubbleplot}
bubble(ma.mr)
```

```{r multivreg}
ma <- metagen(TE=lnes, seTE=selnes, studlab = paste(author, pub_year), sm="OR", comb.fixed = F, data=data[!is.na(data$hpvrate),])
metareg(ma,~hpvrate+aair,method.tau = "REML")
```

If we consider HPV rate and age-adjusted incidence rate in the model together, we find the association between IUD use and cervical cancer varies as a function of the age-adjusted incidence rate adjusting for HPV rate.
```{r sessionInfo}
sessionInfo()
```