---
title: "Network Meta Analysis - Uterine Bleeding"
author: "ks"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## Dataset

```{r libraries}
if(!require(netmeta)) install.packages("netmeta")
library(netmeta)
library(meta)
library(readxl)
```

```{r ReadData}
ubdata <- read_xlsx("~kims/Google Drive/My Drive/Teaching/PM605/2021/week14_apr26/nma-uterine-bleeding.xlsx",sheet="orig-data")
ubdata
```

```{r mbin1}
ma1 <- metabin(t1events, t1n, t2events, t2n, sm="OR", 
               method="MH", data=ubdata[ubdata$treat2=="Placebo",],
               studlab=Study)
ma1
```


```{r mbin2}
ma2 <- metabin(t1events, t1n, t2events, t2n, sm="OR", 
               data=ubdata[ubdata$treat2=="LNG",],
               studlab=Study)
ma2
```

## Network meta-analysis

We have a network meta-analysis for binary data that will use the Mantel-Haenszel method.
```{r netmetabin}
mn1 <- netmetabin(event1=t1events,n1=t1n,
               event2=t2events,n2=t2n, treat1, treat2, Study,
         data=ubdata, sm="OR")
```


```{r graph-it}
netgraph(mn1, seq=c("Placebo", "COCP", "LNG"))
```

```{r print}
print.netmeta(mn1)
```

Summarize output:

```{r summary}
print(summary(mn1))
```


## Placebo Reference

```{r placebo-ref}
summary(mn1, ref="Placebo")
```

Forest Plot for fixed effect model:

```{r placebo-ref-forest}
forest(mn1, ref="Placebo",
       leftlabs="Contrast to Placebo",
       xlab="Odds Ratio",
       pooled="fixed")
```

We can estimate the treatment effect of LNG vs Placebo, but $\it should$ we? 

## combine sORs

What happens if I try to just combine the summary ORs?
```{r ReadData-sORs}
ubdata <- read_xlsx("~kims/Google Drive/My Drive/Teaching/PM605/2021/week14_apr26/nma-uterine-bleeding.xlsx",sheet="sORs")
ubdata
```

```{r get-TEs}
selnes <- with(ubdata,(log(`95UB`)-log(`95LB`))/2/1.96)
esdata <- cbind.data.frame(ubdata,
                           lnes=log(ubdata$OR),
                           selnes)
esdata
```

```{r netmeta}
mn1 <- netmeta(TE=lnes,seTE=selnes, treat1, treat2, Study,
         data=esdata, sm="OR")
print(mn1)
```

It looks like I get the same answer, within rounding error.

## SessionInfo

```{r sessioninfo}
sessionInfo()
```