---
title: "Sparse data meta-analysis"
author: "ks"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## load Libraries 

```{r packages}
for (pkg in c("tidyverse","MASS","purrr", "readxl","dplyr","meta")) 
   library(pkg, character.only = TRUE)

str_detect = stringr::str_detect
```

## Read EFM data 

```{r ReadData}
exdata <- read_xlsx("~kims/Google Drive/My Drive/Teaching/PM605/pm-605-sp2021/week11-apr5/efmstudy.xlsx",sheet=1)
exdata
```

## Default Settings 

What happens if we run a meta-analysis with zero counts in our data set?

```{r mbincc}
mb1 <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="MH", data=exdata, studlab=Studyno)
mb1
```

The output tells us the software used a continuity correction of 0.5 in studies with zero cell frequencies. 

## Treatment Arm weights

Let's try calling the opposite treatment arm sample size weights instead (TACC).

```{r mbintacc}
mbtacc <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="MH", data=exdata, incr = "TACC",
               studlab=Studyno)
mbtacc
```

That made a big difference on the OR estimate from study number 2.

## 'Hand' Calculations

Now let's check the calculation of these OR estimates.

```{r get-sumstats}
ors <- with(exdata, Tevents*(CN-Cevents)/(TN-Tevents)/Cevents)
vares <- with(exdata, 1/Tevents + 1/(TN-Tevents) + 1/Cevents + 1/(CN-Cevents))

lnes <- log(ors)
selnes <- sqrt(vares)
esdata <- cbind.data.frame(Study = exdata$Studyno,
                           ors,
                           lbs = exp(lnes-1.96*selnes),
                           ubs = exp(lnes+1.96*selnes),
                           lnes,
                           selnes)
esdata
```

```{r metabinoutput}
mbtacc
```

Now use a continuity correction based on the inverse treatment sample sizes.
```{r get-zeros}
sstudies <- exdata %>% 
             filter(Tevents==0) %>%
             mutate(R = CN/TN,
                    kt = 1/(R+1),
                    kc = R/(R+1))

ors <- with(sstudies, (Tevents+kt)*(CN-Cevents+kc)/(TN-Tevents+kt)/(Cevents+kc))
vares <- with(sstudies, 1/(Tevents+kt) + 1/(TN-Tevents+kt) + 1/(Cevents+kc) + 1/(CN-Cevents+kc))

lnes <- log(ors)
selnes <- sqrt(vares)
sesdata <- cbind.data.frame(Study = sstudies$Studyno,
                            ors,
                           lbs = exp(lnes-1.959964*selnes),
                           ubs = exp(lnes+1.959964*selnes),
                           lnes,
                           selnes)
sesdata
```

These match the results from calling incr="TACC".  I think the difference in study 2's UB is rounding error.

## Non-0 study OR

What is the MH estimate fromm the studies with non-zero outcomes in the treatment group?

```{r non-zerostudies}
essum <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="MH", data=exdata, 
               incr="TACC",subset= (Tevents>0),
               studlab=Studyno)
essum
```

For the empirical correction, use estimate of 0.3 in continuity correction.

## Empirical CC

I did not find this analysis as an option in the metabin function. We'll have program this one ourselves.

```{r empirical-zstudies}
sstudies <- exdata %>% 
             filter(Tevents==0) %>%
             mutate(R = CN/TN,
                    kt = 0.3/(R+0.3),
                    kc = R/(R+0.3))

ors <- with(sstudies, (Tevents+kt)*(CN-Cevents+kc)/(TN-Tevents+kt)/(Cevents+kc))
vares <- with(sstudies, 1/(Tevents+kt) + 1/(TN-Tevents+kt) + 1/(Cevents+kc) + 1/(CN-Cevents+kc))

lnes <- log(ors)
selnes <- sqrt(vares)
sesdata <- cbind.data.frame(Study = sstudies$Studyno,
                            Tevents = sstudies$Tevents + sstudies$kt,
                            TN = sstudies$TN + 2*sstudies$kt,
                            Cevents = sstudies$Cevents + sstudies$kc,
                            CN = sstudies$CN + 2*sstudies$kc,
                            ors,
                           lbs = exp(lnes-1.96*selnes),
                           ubs = exp(lnes+1.96*selnes),
                           lnes,
                           selnes )
sesdata
```

Combine these data with data from the 'non-zero' studies.
```{r get-ors-nz}
nzstudies <- exdata %>% 
             filter(Tevents>0)

ors <- with(nzstudies, (Tevents)*(CN-Cevents)/(TN-Tevents)/(Cevents))
vares <- with(nzstudies, 1/(Tevents) + 1/(TN-Tevents) + 1/(Cevents) + 1/(CN-Cevents))

lnes <- log(ors)
selnes <- sqrt(vares)
nzesdata <- cbind.data.frame(Study = nzstudies$Studyno,
                             Tevents = nzstudies$Tevents,
                             TN = nzstudies$TN,
                             Cevents = nzstudies$Cevents,
                             CN = nzstudies$CN,
                             ors,
                           lbs = exp(lnes-1.96*selnes),
                           ubs = exp(lnes+1.96*selnes),
                           lnes,
                           selnes)
```

```{r combine}
empiricalcc <- rbind.data.frame(sesdata,nzesdata)
empiricalcc
```


```{r empiricalcc}
essum <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="MH", data=empiricalcc,
               studlab=Study)
essum
```

Nice! The results match the estimates from the Stat Med 2004 paper.

How different is the summary if we use inverse variance weights?
```{r empir-inverse}
essum <- metabin(Tevents, TN, Cevents, CN, sm="OR",
                 method = "Inverse",data=empiricalcc,
               studlab=Study)
essum
```

The results are similar and also match the estimates from the Stat Med 2004 paper.

## SessionInfo

```{r sI}
sessionInfo()
```