---
title: "Sparse pio data meta-analysis"
author: "ks"
date: "4/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# {.tabset}

## load Libraries 

```{r packages}
for (pkg in c("tidyverse","MASS","purrr", "readxl","dplyr","meta")) 
   library(pkg, character.only = TRUE)

str_detect = stringr::str_detect
```

## Read PIO-BLCA data 

```{r ReadData}
exdata <- read_xlsx("data/210404-RCT-data-for-Pio-BLCA.xlsx",sheet=1)
exdata
```

## Flag = P or O 

Analyze the Per Protocol data (flag="P") and the original Dormandy PROactive results (flag="O").

```{r mbincc}
sdata <- exdata %>% 
           filter(is.element(flag,c("P","O")) )
sdata
```

```{r mbinccsum}
mb1 <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="MH", data=sdata, incr = "TACC",
               studlab=PioRCTs)
summary(mb1)
```

## Revised counts

```{r revcounts}

exdata <- read_xlsx("data/210404-RCT-data-for-Pio-BLCA.xlsx",sheet=1)
sdata <- exdata %>% 
           filter(is.element(flag,c("P","R")) )
sdata
```

```{r }
mbtacc <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="MH", data=sdata, incr = "TACC",
               studlab=PioRCTs)
summary(mbtacc)
```

```{r forest}
forest(mbtacc)
```


## Cumulative Meta-analysis

Cumulative meta-analysis by calendar year.
Disappointing, but it looks like the figure we created earlier can only be done for linear models, and in our case inverse variance weighting instead of MH summary.
Before I said the MH OR is preferred. How different are our results if we use the inverse variance weights?
```{r }
mbtacc <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="Inverse", 
               data=sdata, incr = "TACC",
               studlab=PioRCTs)
summary(mbtacc)
```

They don't give me any different results.  I'll use this for the cummulative meta-analysis, instead of programming something from scratch just because it's easier and it doesn't seem like our results are sensitive to the model choice for these data.

```{r cummeta}
library('metafor')
res <- rma(yi=mbtacc$TE, 
           sei=mbtacc$seTE)
 
### cumulative meta-analysis (in the order of seTE)
tmp <- cumul(res, order=order(sdata$calendar.order))
### cumulative forest plot
forest(tmp, xlim=c(-1,1), at=log(c(0.4, 1, 15)), 
       atransf=exp, digits=c(2,3), cex=0.5)
```    


## Non-zero Summary OR

Let's look at the studies with non-zero events only.
```{r nonzero}
cdata <- sdata %>%
          filter(Tevents>0 & Cevents >0)
mbtacc <- metabin(Tevents, TN, Cevents, CN, sm="OR", 
               method="MH", data=cdata, 
               studlab=PioRCTs)
summary(mbtacc)
```

This result looks much the same, which is no surprise given that these 2 studies had >85% of the weight in the analysis of 5 studies.

## SessionInfo

```{r sI}
sessionInfo()
```